---
title: "HW 3, pt. 2: Exploring U.S. Supreme Court Decisions"
author: "Julia Du"
output: github_document
---

## Get the data

```{r}
# load useful packages
library(tidyverse)
library(readr)
library(lubridate)
library(ggthemes)
library('scales')
library(forcats)

# load data
scdb_case <- read_csv("data/scdb-case.csv")
#View(scdb_case)
#, col_types = cols(sctCite = col_character())
#can docket be a character? soln for this 1st part leaves sctCite alone as a logical object

scdb_vote <- read_csv("data/scdb-vote.csv")
#View(scdb_vote)
```

## Recode variables as you find necessary

```{r}
#does this mean parsing? do i only need to parse the variables that i'll be working w/?
#also, i can't just drop the variables not relevant to HW 3, right?

#term right now is a double...is that okay...

```

## What percentage of cases in each term are decided by a one-vote margin (i.e. 5-4, 4-3, etc.)

```{r}
scdb_case %>%
  drop_na(term, majVotes, minVotes) %>%
  group_by(term) %>%
  mutate(V_margin = majVotes-minVotes) %>%
  mutate(onevotelgl = (V_margin == 1)) %>%
  summarize(
    totalcases = n(),
    onevote = mean(onevotelgl)
  ) %>%
 ggplot(mapping = aes(x = term, y = onevote)) +
  geom_line() +
  labs(title = "Percent of U.S. Supreme Court cases decided by 1-vote margin", x = "Term", y = "Percent of total cases decided", caption = "Source: The Supreme Court Database") +
 scale_y_continuous(labels = scales::percent) +
  theme_solarized()
```

## For justices [currently serving on the Supreme Court](https://www.supremecourt.gov/about/biographies.aspx), how often have they voted in the conservative direction in cases involving criminal procedure, civil rights, economic activity, and federal taxation?

Organize the resulting graph by justice in descending order of seniority. (Note that the chief justice is always considered the most senior member of the court, regardless of appointment date.)

```{r}
require(forcats)

issueArea_names <- c(
                    `1` = "Criminal Procedure",
                    `2` = "Civil Rights",
                    `8` = "Economic Activity",
                    `12` = "Federal Taxation"
                                        )
justice_levels = c("JGRoberts", "CThomas", "SGBreyer", "SAAlito", "SSotomayor", "EKagan", "NMGorsuch", "BMKavanaugh", "ACBarrett") 

#idk idk
scdb_vote %>%
  mutate(alljust = distinct(justiceName)) %>%
  mutate(seniorjust = factor(levels = justice_levels))
#  mutate(currentjust = 


scdb_case %>%
  drop_na(issueArea) %>%
  group_by(issueArea) %>%
  filter(
    issueArea %in% c(1, 2, 8, 12)
    ) %>%
  select(caseIssuesId, issueArea) %>%
   right_join(scdb_vote, by = "caseIssuesId") %>%
  group_by(justiceName, issueArea) %>%
   filter(
    justiceName %in% c("JGRoberts", "CThomas", "SGBreyer", "SAAlito", "SSotomayor", "EKagan", "NMGorsuch", "BMKavanaugh", "ACBarrett")) %>%
  drop_na(direction, issueArea) %>%  ##CAREFUL -- check on why you're getting NAs from issueArea
  mutate(conservlgl = (direction == 1)) %>%
  summarize(
    totalvotes = n(),
    conserv = mean(conservlgl)
  ) %>%
  ggplot(mapping = aes(x = conserv, y = reorder(justiceName, seniorjust))) + ###FIGURE OUT HOW TO ORDER BY SENIORITY
  geom_bar(stat = 'identity') +
  labs(title = "US Supreme Court", subtitle = "Percent of cases decided in a conservative direction", x = "Percent of votes cast", y = " ", caption = "Source: The Supreme Court Database") +
 scale_x_continuous(labels = scales::percent) +
  facet_wrap(issueArea ~ ., labeller = as_labeller(issueArea_names)) +
  theme_solarized()


#or maybe only grouping by issues after all & then using facet(JusticeName)? would that mean i have to undo the summarize bit?
ggplot(mapping = aes(x = conserv, y = issueArea)) + ###FIGURE OUT HOW TO ORDER BY SENIORITY
  geom_bar(stat = 'identity') +
  labs(title = "US Supreme Court", subtitle = "Percent of cases decided in a conservative direction", x = "Percent of votes cast", caption = "Source: The Supreme Court Database") +
 scale_x_continuous(labels = scales::percent) +
  facet_wrap(justiceName ~ ., labeller = as_labeller(issueArea_names)) +
  theme_solarized()
#this above was my attempt at making a graph that was faceted by justice - it didn't work. maybe cuz i'd have to summarize in a diff way (i.e. issueArea, then justiceName). do i have to make all these graphs

#may want to create factor of justices 
# use levels to sort justices by descending seniority
scdb_vote %>%
  group_by(justiceName) %>%
  filter(
    justiceName %in% c("JGRoberts", "CThomas", "SGBreyer", "SAAlito", "SSotomayor", "EKagan", "NMGorsuch", "BMKavanaugh", "ACBarrett")) %>%
  #have to join this to the scdb_case data
  #filter(issueArea %in% c(1, 2, 8, 12))
```

## In each term, how many of the term's published decisions (decided after oral arguments) were announced in a given month?

```{r}

```

## Which justices are most likely to agree with with the Court's declaration that an act of Congress, a state or territorial law, or a municipal ordinance is unconstitutional? Identify all cases where the Court declared something unconstitutional and determine the ten justices who most and least frequently agreed with this outcome as a percentage of all votes cast by the justice in these cases. Exclude any justice with fewer than 30 votes in cases where the Court's outcome declares something unconstitutional.

```{r}

```

## In each term he served on the Court, in what percentage of cases was Justice Antonin Scalia in the majority?

```{r}

```

## Create a graph similar to above that adds a second component which compares the percentage for all cases versus non-unanimous cases (i.e. there was at least one dissenting vote)

```{r}

```

## In each term, what percentage of cases were decided in the conservative direction?

```{r}

```

## The Chief Justice is frequently seen as capable of influencing the ideological direction of the Court. Create a graph similar to the one above that also incorporates information on who was the Chief Justice during the term.

```{r}

```

## Session info

```{r, echo = TRUE}
devtools::session_info()
```


