---
title: "HW 3, pt. 2: Exploring U.S. Supreme Court Decisions"
author: "Julia Du"
output: github_document
---

## Get the data

```{r}
# load useful packages
library(tidyverse)
library(readr)
library(lubridate)
library(ggthemes)
library('scales')
library(forcats)

theme_set(theme_solarized())

# load data
scdb_case <- read_csv("data/scdb-case.csv")
scdb_case

scdb_vote <- read_csv("data/scdb-vote.csv")
scdb_vote
```

## Recode variables as you find necessary
I decided not to recode variables at the beginning. If I needed to recode anything for a question, I did it within the assigned code chunk.
```{r}

```

## What percentage of cases in each term are decided by a one-vote margin (i.e. 5-4, 4-3, etc.)
**Note:** Whenever I mutate a new variable with "lgl" as a suffix, that's a note to myself that I'm creating a logical variable (which I then use to find a %). For most of this HW, I prefer mutating a new logical var instead of mutating a recoded var, but I do mutate a recoded variable in Q8 to show that these 2 methods give me the same answer.

**Also,** throughout my mutation of logical/recoded variables in this HW, I often summarize to find the total counts alongside the mean (i.e. the percentages). For most questions, the total counts aren't needed for the answer, but I've kept them in because they are useful in checking my work.

```{r}
scdb_case %>%
  drop_na(term, majVotes, minVotes) %>%
  group_by(term) %>%
  mutate(V_margin = majVotes-minVotes) %>%
  mutate(onevotelgl = (V_margin == 1)) %>%
  summarize(
   # totalcases = n(),
    onevote = mean(onevotelgl)
  ) %>%
 ggplot(mapping = aes(x = term, y = onevote)) +
  geom_line() +
  labs(title = "Percent of U.S. Supreme Court cases decided by 1-vote margin", x = "Term", y = "Percent of total cases decided", caption = "Source: The Supreme Court Database") +
 scale_y_continuous(labels = scales::percent) 
```

## For justices [currently serving on the Supreme Court](https://www.supremecourt.gov/about/biographies.aspx), how often have they voted in the conservative direction in cases involving criminal procedure, civil rights, economic activity, and federal taxation?

Organize the resulting graph by justice in descending order of seniority. (Note that the chief justice is always considered the most senior member of the court, regardless of appointment date.)

```{r}
require(forcats)

issueArea_names <- c(
                    `1` = "Criminal Procedure",
                    `2` = "Civil Rights",
                    `8` = "Economic Activity",
                    `12` = "Federal Taxation"
                                        )
justice_levels <- c("JGRoberts", "CThomas", "SGBreyer", "SAAlito", "SSotomayor", "EKagan", "NMGorsuch", "BMKavanaugh") 
#This dataset is for the 1791-2019 SCOTUS terms, which each start in October. 
#Though she is currently serving, Amy Coney Barrett isn't in this dataset (she was sworn in Oct. 2020) so I left her out.

#facet by issueArea 
#note: this method is not as efficient as my "facet by justiceName" chunk below, but I wanted to keep this code since it goes through my thought process step-by-step
scdb_case %>%
  drop_na(issueArea) %>%
  group_by(issueArea) %>%
  filter(
    issueArea %in% c(1, 2, 8, 12)
    ) %>%
  select(caseIssuesId, issueArea) %>%
   right_join(scdb_vote, by = "caseIssuesId") %>%
mutate(OrderjusticeName = fct_relevel(justiceName, justice_levels)) %>% 
    group_by(OrderjusticeName, issueArea) %>%
   filter(
    OrderjusticeName %in% c("JGRoberts", "CThomas", "SGBreyer", "SAAlito", "SSotomayor", "EKagan", "NMGorsuch", "BMKavanaugh")
    ) %>%
  drop_na(direction, issueArea) %>%  
      mutate(conservlgl = (direction == 1)) %>%
  summarize(
    totalvotes = n(),
    conserv = mean(conservlgl)
  ) %>%
  ggplot(mapping = aes(x = conserv, y = fct_rev(OrderjusticeName))) + 
  geom_bar(stat = 'identity') +
  labs(title = "US Supreme Court", subtitle = "Percent of cases decided in a conservative direction", x = "Percent of votes cast", y = " ", caption = "Source: The Supreme Court Database") +
 scale_x_continuous(labels = scales::percent) +
  facet_wrap(issueArea ~ ., labeller = as_labeller(issueArea_names)) 

#facet by justice
scdb_case %>%
  select(caseIssuesId, issueArea) %>%
   right_join(scdb_vote, by = "caseIssuesId") %>%
mutate(OrderjusticeName = fct_relevel(justiceName, justice_levels)) %>% ###delete if this doesn't work
  drop_na(direction, issueArea, OrderjusticeName) %>%  
     filter(
    issueArea %in% c(1, 2, 8, 12),
    OrderjusticeName %in% c("JGRoberts", "CThomas", "SGBreyer", "SAAlito", "SSotomayor", "EKagan", "NMGorsuch", "BMKavanaugh")
    ) %>%
  group_by(OrderjusticeName, issueArea) %>%
      mutate(conservlgl = (direction == 1)) %>%
  summarize(
    conserv = mean(conservlgl)
  ) %>%
 mutate(issueArea = factor(issueArea, labels = issueArea_names)) %>%
  ggplot(mapping = aes(x = conserv, y = fct_rev(issueArea))) +
  geom_col() +
  labs(title = "US Supreme Court", subtitle = "Percent of cases decided in a conservative direction", x = "Percent of votes cast", y = "", caption = "Source: The Supreme Court Database") +
 scale_x_continuous(labels = scales::percent) +
  facet_wrap(OrderjusticeName ~ .) 
```

## In each term, how many of the term's published decisions (decided after oral arguments) were announced in a given month?

```{r}
trymonth = c(` ` = month.name) 

scdb_case %>%
  mutate(date_parsed = mdy(dateDecision)) %>%
  mutate(month_announce = month(date_parsed)) %>%
  mutate(month_announce = factor(month_announce, labels = trymonth)) %>%
    drop_na(term, month_announce, decisionType) %>%
#  select(term, month_announce, decisionType) %>%
  filter(decisionType %in% c(1, 6, 7)) %>% #filtering only for cases that were orally argued
  group_by(term, month_announce) %>%
    count(month_announce) %>%
  mutate(month_announce = factor(month_announce, levels = month.name)) %>%
    mutate(month_announce = fct_shift(month_announce, n = -3)) %>%
 ggplot() +
    geom_boxplot(mapping = aes(x = n, y = fct_rev(month_announce))) +
    labs(title = "U.S. Supreme Court", subtitle = "Number of decisions announced post-oral arguments per month, by term", x = "Number of decisions announced in a term-month", y = "", caption = "Source: The Supreme Court Database") 
```

## Which justices are most likely to agree with with the Court's declaration that an act of Congress, a state or territorial law, or a municipal ordinance is unconstitutional? Identify all cases where the Court declared something unconstitutional and determine the ten justices who most and least frequently agreed with this outcome as a percentage of all votes cast by the justice in these cases. Exclude any justice with fewer than 30 votes in cases where the Court's outcome declares something unconstitutional.

```{r}
scdb_case %>%
  group_by(declarationUncon) %>%
  filter(
    declarationUncon > 1
    ) %>%
  select(caseIssuesId, declarationUncon) %>%
   right_join(scdb_vote, by = "caseIssuesId") %>%  
   group_by(justiceName) %>%
  drop_na(majority, justiceName, declarationUncon) %>%  
  mutate(yes_unconlgl = (majority == 2)) %>%
  summarize(
    vote_uncon = n(),
    yes_uncon = mean(yes_unconlgl)
  ) %>%
  filter(vote_uncon >= 30) %>%
  arrange(-yes_uncon) %>%
  slice(1:10, 71:80) %>%
  mutate(type = ifelse(yes_uncon>0.90, "top", "bottom")) %>% 
  ggplot(mapping = aes(x = yes_uncon, y = fct_reorder(justiceName, .x = yes_uncon), fill = type)) +
  geom_col() +
  scale_x_continuous(labels = scales::percent) +
  labs(title = "U.S. Supreme Court", subtitle = "Agreement with a declaration of unconstitutionality \n Justices most and least frequently in agreement", x = "Percent of votes agreeing with declaration", y = "", caption = "Source: The Supreme Court Database") +
  theme_bw() +
    theme(legend.position = "none")
```

## In each term he served on the Court, in what percentage of cases was Justice Antonin Scalia in the majority?

```{r}
scdb_vote %>%
    filter(
    justiceName == "AScalia"
    ) %>%
  group_by(term) %>%
  drop_na(majority, justiceName, term) %>%  
  mutate(Scalia_majlgl = (majority == 2)) %>%
  summarize(
    totalScalia_maj = n(),
    Scalia_maj = mean(Scalia_majlgl)
  ) %>%
 ggplot(mapping = aes(x = term, y = Scalia_maj)) +
  geom_line() +
  labs(title = "Percent of the time Scalia was in the majority", x = "Term", y = "Percent of total decisions", caption = "Source: The Supreme Court Database") +
 scale_y_continuous(labels = scales::percent) 
```

## Create a graph similar to above that adds a second component which compares the percentage for all cases versus non-unanimous cases (i.e. there was at least one dissenting vote)

```{r}
alldecisions <- scdb_vote %>%
    filter(
    justiceName == "AScalia"
    ) %>%
  group_by(term) %>%
  drop_na(majority, justiceName, term) %>%  
  mutate(Scalia_majlgl = (majority == 2)) %>%
  summarize(
    totalScalia_maj = n(),
    Scalia_maj = mean(Scalia_majlgl)
  )

nonall_decisions <-  
  scdb_case %>%
  select(caseIssuesId, minVotes) %>%
  right_join(scdb_vote, by = "caseIssuesId") %>%
  drop_na(minVotes, majority, justiceName, term) %>%
  filter(
    minVotes != "0",
    justiceName == "AScalia"
    ) %>%
  group_by(term) %>%
  mutate(dis_Scalia_majlgl = (majority == 2)) %>%
  summarize(
#    dis_totalScalia_maj = n(),
    dis_Scalia_maj = mean(dis_Scalia_majlgl)
  ) 

#plotting w/ 2 dataframes
Scaliacolors <- c("All decisions" = "blue", "Non-unanimous decisions" = "red")

ggplot() +
  geom_line(data = alldecisions, mapping = aes(x = term, y = Scalia_maj, color = "All decisions")) +
  geom_line(data = nonall_decisions, mapping = aes(x = term, y = dis_Scalia_maj, color = "Non-unanimous decisions")) +
  labs(title = "Percent of the time Scalia was in the majority", x = "Term", y = "Percent of total decisions", caption = "Source: The Supreme Court Database", color = "Legend") +
  scale_color_manual(values = Scaliacolors) +
 scale_y_continuous(labels = scales::percent) +
  theme_bw() +
  theme(legend.title = element_blank()) +
  theme(legend.position="bottom")

#plotting w/ 1 dataframe
alldecisions %>%
  inner_join(nonall_decisions, by = "term") %>%
  pivot_longer(cols = c(Scalia_maj, dis_Scalia_maj), names_to = "type", values_to = "percentage") %>%
ggplot(mapping = aes(x = term, y = percentage, color = type)) +
  geom_line() +
  labs(title = "Percent of the time Scalia was in the majority", x = "Term", y = "Percent of total decisions", caption = "Source: The Supreme Court Database", color = "Legend") +
 scale_y_continuous(labels = scales::percent) +
  scale_color_discrete(labels = c("Non-unanimous decisions", "All decisions")) +
   theme_bw() +
  theme(legend.title = element_blank()) +
  theme(legend.position="bottom") 
```

## In each term, what percentage of cases were decided in the conservative direction?

```{r}
scdb_case %>%
  drop_na(term, decisionDirection) %>%
  group_by(term) %>%
  mutate(conserv_caselgl = (decisionDirection == 1)) %>%
  summarize(
 #   totalconserv_cases = n(),
    conserv_case = mean(conserv_caselgl)
  ) %>%
 ggplot(mapping = aes(x = term, y = conserv_case)) +
  geom_line() +
  labs(title = "U.S. Supreme Court", subtitle = "Percent of cases decided in a conservative direction", x = "Term", y = "Percent of total decisions", caption = "Source: The Supreme Court Database") +
 scale_y_continuous(labels = scales::percent) 
```

## The Chief Justice is frequently seen as capable of influencing the ideological direction of the Court. Create a graph similar to the one above that also incorporates information on who was the Chief Justice during the term.
**Note:** This time, to find % for the basic facet graph (i.e. you ony see a justice's individual term), I mutated a recoded variable instead of mutating a new logical variable (which has been my preferred method throughout this HW). 

*I can still find % with a logical variable* (& I do so under the trychief_facet code chunk, when finding the facet graph with opaque all terms), but I wanted to show that I can use this method as well - and that these 2 methods give me the same answer.

```{r}
chief_levels = c("Jay", "Rutledge", "Ellsworth", "Marshall", "Taney", "Chase", "Waite", "Fuller", "White", "Taft", "Hughes", "Stone", "Vinson", "Warren", "Burger", "Rehnquist", "Roberts") 

#basic graph (only see justice's individual terms)
scdb_case %>%
  mutate(Orderchief = fct_relevel(chief, chief_levels)) %>%
  drop_na(term, decisionDirection, Orderchief) %>%
   group_by(term, Orderchief) %>%
    mutate(Recode_decDirect = ifelse(decisionDirection != "1", 0, 1)) %>% #do this != to ensure we account for unspecified
    summarize(
#    totalconserv_cases = n(),
    conserv_case = mean(Recode_decDirect, na.rm = TRUE)
  ) %>%
  ggplot(mapping = aes(x = term, y = conserv_case)) +
  geom_line() +
 facet_wrap(Orderchief ~ .) +
  labs(title = "U.S. Supreme Court", subtitle = "Percent of cases decided in a conservative direction", x = "Term", y = "Percent of total decisions", caption = "Source: The Supreme Court Database") +
 scale_y_continuous(labels = scales::percent) +
  theme_bw()

#facet graph (with opaque all terms)
fullchief_facet <- scdb_case %>%
  drop_na(term, decisionDirection) %>%
  group_by(term, chief) %>% #we keep this as chief & don't switch to Orderchief so that faceting by Orderchief works later
  mutate(conserv_caselgl = (decisionDirection == 1)) %>%
  summarize(
#    totalconserv_cases = n(),
    conserv_case1 = mean(conserv_caselgl)
  ) 
trychief_facet <- scdb_case %>%
  mutate(Orderchief = fct_relevel(chief, chief_levels)) %>%
  drop_na(term, decisionDirection, Orderchief) %>%
   group_by(term, Orderchief) %>%
  mutate(conserv_caselgl = (decisionDirection == 1)) %>%
  summarize(
#    totalconserv_cases = n(),
    conserv_case = mean(conserv_caselgl, na.rm = TRUE)
  )
ggplot() +
 geom_line(data = fullchief_facet, mapping = aes(x = term, y = conserv_case1), alpha = 0.1) +
    geom_line(data = trychief_facet, mapping = aes(x = term, y = conserv_case), color = "red") +
  facet_wrap(Orderchief ~ .) +
 labs(title = "U.S. Supreme Court", subtitle = "Percent of cases decided in a conservative direction", x = "Term", y = "Percent of total decisions", caption = "Source: The Supreme Court Database") +
   scale_y_continuous(labels = scales::percent) +
  theme_bw() +
  theme(legend.position = "none")

#chief shade graph
##to get % for all terms
trychief_facetline <- trychief_facet %>% 
  ggplot(mapping = aes(x = term, y = conserv_case)) +
  geom_line() +
  labs(title = "U.S. Supreme Court", subtitle = "Percent of cases decided in a conservative direction", x = "Term", y = "Percent of total decisions", caption = "Source: The Supreme Court Database") +
 scale_y_continuous(labels = scales::percent) 

##to get shading by Chief Justice
chief_facet <- scdb_case %>%
 mutate(Orderchief = fct_relevel(chief, chief_levels)) %>%
  distinct(term, Orderchief) %>%
  group_by(Orderchief) %>%
  mutate(xmin = min(term), xmax = max(term)) %>%
  select(-term) %>%
  distinct() %>%
  mutate(ymin = min(trychief_facet$conserv_case), ymax = max(trychief_facet$conserv_case))

trychief_facetline + 
  geom_rect(data = chief_facet, mapping = aes(NULL, NULL, xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = Orderchief), alpha = 0.2) +
  labs(fill = "Chief Justice") +
  theme_bw()
```

## Session info

```{r, echo = TRUE}
devtools::session_info()
```


